---
title       : Chance of Esophageal Cancer
subtitle    : How likely are you to have Esophageal Cancer?
author      : Kenneth Kihara
job         : Programmer
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Usage
This shiny app calculates the chance that a person will have esophageal 
cancer given age, alcohol consumption, and tobacco consumption. Choose 
the inputs by moving the appropriate slider. The output is printed under 
"Predicted Chance of Esophageal Cancer:"

## Data
The data used for predicting esophageal cancer was taken from the esoph 
dataset in the R datasets package. The predictors used are:

- Age Group
- Daily Alcohol Consumption
- Daily Tobacco Consumption

---

## Why use this application?

1. Promote awareness of esophageal cancer
2. Diagnose and seek treatment sooner
3. Supporting evidence that alchol and tobacco causes cancer
4. Determine a quick prediction of your chance of cancer
5. Show the dangers of alcohol and tobacco consumption

---

## Accuracy
The model was trained with a 10-fold cross validation. Giving an accuracy 
and standard deviation of:
```{r, echo=FALSE}
library(caret)
library(datasets)
df <- esoph
x <- factor(esoph$agegp, levels=levels(esoph$agegp), 
            labels=c("29.5","39.5","49.5","59.5","69.5","79.5"))
df$agegp <- as.numeric(as.character(x))
x <- factor(esoph$alcgp, levels=levels(esoph$alcgp), 
            labels=c("19.5","59.5","99.5","139.5"))
df$alcgp <- as.numeric(as.character(x))
x <- factor(esoph$tobgp, levels=levels(esoph$tobgp), 
            labels=c("4.5","14.5","24.5","34.5"))
df$tobgp <- as.numeric(as.character(x))
fin_data <- NULL
for (i in 1:nrow(esoph)) {
    num_cases <- df$ncases[i]
    num_ctrls <- df$ncontrols[i] - num_cases
    cases <- NULL
    ctrls <- NULL
    row <- c(df$agegp[i], df$alcgp[i], df$tobgp[i])
    while (num_cases > 0) {
        cases <- rbind(cases, c(row, 1))
        num_cases <- num_cases - 1
    }
    while (num_ctrls > 0) {
        ctrls <- rbind(ctrls, c(row, 0))
        num_ctrls <- num_ctrls - 1
    }
    fin_data <- rbind(fin_data, cases, ctrls)
}
fin_data <- data.frame(fin_data)
names(fin_data) <- c("agegp","alcgp","tobgp","outcome")
ctrl <- trainControl(method="repeatedcv", number=10, repeats=10)
fit <- train(factor(outcome)~.,data=fin_data,method="glm",family="binomial",
             trControl=ctrl)
```

```{r}
fit$results$Accuracy
fit$results$AccuracySD
```

---

## Plots

Plots of frequency of age vs alcohol consumption and age vs tobacco 
consumption.
```{r, echo=FALSE}
pos <- subset(fin_data, outcome==1)
pos$n <- 1
agevsalc <- aggregate(n~agegp+alcgp, data=pos, FUN=sum)
agevstob <- aggregate(n~agegp+tobgp, data=pos, FUN=sum)

par(mfrow=c(1,2))
plot(agevsalc$agegp, agevsalc$alcgp, type="n",
     xlab="Age (years)", ylab="Alcohol Consumption (gm/day)",
     main="Alcohol vs Age Frequencies")
points(agevsalc$agegp, agevsalc$alcgp, pch=21, cex=agevsalc$n/5, 
       bg="steelblue2")
plot(agevstob$agegp, agevstob$tobgp, type="n",
     xlab="Age (years)", ylab="Tobacco Consumption (gm/day)",
     main="Tobacco vs Age Frequencies")
points(agevstob$agegp, agevstob$tobgp, pch=21, cex=agevstob$n/5, 
       bg="steelblue2")
```
